Response to reviewers

### 2025-08-20 src/docs_paper-jtb/revisions.Rmd
JIE suggestions (#61)
* JIE suggestions

* Add cite

* Integrate JIE changes

---------

Co-authored-by: athowes <adamthowes@gmail.com>

Dear Dr. XXXX,

@@

Below we address the reviewers' comments and suggestions point by point. While updating and extending the simulations to address some of the reviewers' comments, we noticed that the results reported arose from two different versions of the model, with some from an alternative experimental algorithm that was not retained for the final method. Along with updating all results with longer MCMC chain simulations, we have now corrected the results in the manuscript to all be from the PCA grid described. This change impacts numerical results reported for the KS test, MMD, and exceedance probabilities, but does not change the overall conclusions or recommendations from the paper.

@@

@@

\textbf{Response:} `tmbstan` passes the objective function and its gradient (calculated in `TMB`) directly to `rstan` via the [`op_matrix_vari`](https://mc-stan.org/math/classstan_1_1math_1_1op__matrix__vari.html).
As such, the NUTS algorithm used is exactly the same (as the same underlying `rstan` code is called.
It could be the case that there are speed differences between function evaluations in `TMB` and Stan.
We have added a sentence to the discussion noting this possible comparison:

@@

\textbf{Response:} Some HIV surveys, including the example data from Malawi used in the analysis reported here, provide bio-marker data on whether or not HIV infection was recently acquired (approximately within the previous 4 months; the $\Omega_T$ parameter in Equation 7). This information on the proportion recently infected along with characteristics of the assay used to measure recent infection provides a direct survey estimator for HIV incidence, as derived by Kassanjee \textit{et al.} (https://pubmed.ncbi.nlm.nih.gov/22627902/). The formula in Equation 7 is a reformulation of the Kassanjee \textit{et al.} estimator to express a likelihood for the survey observations of the number recenlty infected, enabling Bayesian inference on the $\lambda_i$ parameters (HIV incidence rate). Details of the derivation of Equation 7 are described in the supplementary materials of Eaton \text{et al.} (and also Bao \text{et al.} (https://pubmed.ncbi.nlm.nih.gov/25406754/)). We have revised the manuscript here to clarify that Equation 7 is derived from the classical Kassanjee \textit{et al.} estimator:

> The proportion recently infected $\kappa_i$ is linked to HIV incidence via a reformulation of the Kassanjee \textit{et al.} biomarker-based incidence estimator...

@@

\textbf{Response:} In the current work we focused on results which were as accurate as possible while still running in a practical amount of time.
We agree that it is possible that faster results that are nearly as accurate could be obtained using or procedure.

>>>>

Dear Dr. Chaplain,

@@

@@

While updating and extending the simulations to address some of the reviewers' comments, we noticed that the results reported arose from two different versions of the method, with some from an alternative experimental algorithm that was not retained for the final method.
Along with updating all results with longer MCMC chain simulations, we have now corrected the results in the manuscript to all be from the PCA grid described.
This change impacts numerical results reported for the KS test, MMD, and exceedance probabilities, but does not change the overall conclusions or recommendations from the paper.

@@

\textbf{Response:} `tmbstan` passes the objective function and its gradient (calculated in `TMB`) directly to `rstan` via the [`op_matrix_vari`](https://mc-stan.org/math/classstan_1_1math_1_1op__matrix__vari.html).
As such, the NUTS algorithm used is exactly the same (as the same underlying `rstan` code is called).
It could be the case that there are speed differences between function evaluations in `TMB` and Stan.
We have added a sentence to the discussion noting this possible comparison:

@@

\textbf{Response:} Some HIV surveys, including the example data from Malawi used in the analysis reported here, provide bio-marker data on whether or not HIV infection was recently acquired (approximately within the previous 4 months; the $\Omega_T$ parameter in Equation 7). This information on the proportion recently infected along with characteristics of the assay used to measure recent infection provides a direct survey estimator for HIV incidence, as derived by @kassanjee2012new. The formula in Equation 7 is a reformulation of the @kassanjee2012new estimator to express a likelihood for the survey observations of the number recently infected, enabling Bayesian inference on the $\lambda_i$ parameters (HIV incidence rate). Details of the derivation of Equation 7 are described in the supplementary materials of @eaton2021naomi (and also @bao2014incorporating). We have revised the manuscript here to clarify that Equation 7 is derived from the classical @kassanjee2012new estimator:

> The proportion recently infected $\kappa_i$ is linked to HIV incidence via a reformulation of the biomarker-based incidence estimator of @kassanjee2012new...

@@

\textbf{Response:} In the current work we focused on results which were as accurate as possible while still running in a practical amount of time.
We agree that it is possible that faster results that are nearly as accurate could be obtained using our procedure.
